## Human Activity Recognition
========================================================


```{r}
library(rpart)
library(caret)
pml.training <- read.csv("C:/coursera/machine_Learning/Project/pml-training.csv")
pml.testing <- read.csv("C:/coursera/machine_Learning/Project/pml-testing.csv")
```
# Keep only the variables associated with accelerometers on the belt, forearm, arm, and dumbell.  
```{r}
training_pml <- pml.training[,c('accel_belt_x','accel_belt_y','accel_belt_z','accel_arm_x','accel_arm_y','accel_arm_z','accel_dumbbell_x','accel_dumbbell_y','accel_dumbbell_z','accel_forearm_x','accel_forearm_y','accel_forearm_z','classe')]
testing_pml <- pml.testing[,c('accel_belt_x','accel_belt_y','accel_belt_z','accel_arm_x','accel_arm_y','accel_arm_z','accel_dumbbell_x','accel_dumbbell_y','accel_dumbbell_z','accel_forearm_x','accel_forearm_y','accel_forearm_z','problem_id')]
```
# In order to minimise number of predictors Test for highlly correlated variables .
```{r}
corr_pml <- cor(training_pml[, -13])
highCorr <- findCorrelation(corr_pml, 0.90)
highCorr
training_pml_uncrl <- training_pml[,-highCorr ]
testing_pml_uncrl <- testing_pml[, -highCorr]
```
# Inspecting model based on rpart function
```{r}
summary (training_pml)
tc <- trainControl("cv",10)
rpart.grid <- expand.grid(.cp=0.2)
modfit_pm_rpart <- train(classe ~.,method = 'rpart', data = training_pml_uncrl,trControl=tc,tuneGrid=rpart.grid)

modfit_pm_rpart
```
# The rpart model has very low accuracy. Hence it cannot be used for prediction.

Inspect rf model on small sample of training data 20%, due mainlly to memory limit.
```{r}
inTrain_pml_2 <- createDataPartition(y=training_pml_uncrl$classe, p=0.2, list=FALSE)
training_pml_2  <- training_pml_uncrl[inTrain_pml_2,]

modFitRF_pml_2 <- train(classe~., data= training_pml_2, method ="rf", prox=TRUE)
modFitRF_pml_2
```
# Sample validation data for confusion matrix evaluation.  
# In order to avoid using same observation in training and Validation I only used the data that was not used for training
```{r}
valid_pml  <- training_pml_uncrl[-inTrain_pml_2,]

#the data in the validation size is limited to 0.8*0.1 = 0.08 % of the original trining data set
invalid_pml_2 <- createDataPartition(y=valid_pml$classe, p=0.1, list=FALSE)
valid_pml_2 <- valid_pml[invalid_pml_2,]
```
# Inspect the confusion Matrix.
```{r}
valid_pml_2$Prediction <- predict(modFitRF_pml_2, newdata=valid_pml_2)
confusionMatrix(data=valid_pml_2$Prediction, valid_pml_2$classe)
```
# Aplly the predictor model to new data given in the test dataset
```{r}
pred_pml <- predict(modFitRF_pml_2,testing_pml_uncrl )
pred_pml
```
# Attach predicted values to the test dataset
```{r}
testing_pml_uncrl_2 <- testing_pml_uncrl
testing_pml_uncrl_2$pred_class <- pred_pml
#view newe predicted value and row identifier. Needed for the submission part of the course project assignement
testing_pml_uncrl_2[c(12,13)]
```

```{r}

```

